---
title: "In-Class Exercise 3"
author: "Dexter Wan"
date: "September 2, 2024"
date-modified: "last-modified"
execute: 
  eval: true
  echo: true
  freeze: true
---

## Description

Learning some error in Hands-on Exercise 3 and how to solve it.

### Changelog

02 Sep 24: Created In-Class Exercise 3.

## Importing data

Install *maptools.* It has been retired, so we must go to posit public package manager to find the old version.

```{r}
#| eval: false
# The above "eval: false" prevents the code chunk from running everytime the document is rendered
install.packages("maptools", repos = "https://packagemanager.posit.co/cran/2023-10-13")
```

*Maptools* allows us to perform the conversion of *as.SpatialGridDataFrame.im().*

Import Myanmar ACLED data:

```{r}
pacman::p_load(sf, tidyverse)
acled_sf = read_csv("data/ACLED_Myanmar.csv") %>%
  st_as_sf(coords = c("longitude", "latitude"), crs = 4326) %>%
  st_transform(crs = 32647) %>%
  mutate(event_date = dmy(event_date))
```

The read_csv would read the date given as a string. As such, we need to use the mutate command to convert it into a date format.
