---
title: "Hands-on Exercise 6"
author: "Dexter Wan"
date: "September 16, 2024"
date-modified: "last-modified"
execute: 
  eval: true
  echo: true
  freeze: true
---

## Description

Learning to compute Global and Local Measures of Spatial Autocorrelation in R.

### Changelog

16 Sep 24: Created Hands-on Exercise 6.

18 Sep 24: Completed Hands-on Exercise 6.

## Importing Data and setting up R Environment

Loading in relevant packages:

```{r}
pacman::p_load(sf, tidyverse, tmap, spdep)
```

Loading in data:

```{r}
hunan_sf = st_read(dsn = "data/geospatial", layer = "Hunan")
hunan_2012 = read.csv("data/aspatial/Hunan_2012.csv")
hunan_sf = left_join(hunan_sf, hunan_2012)
```

Reduce the columns to only relevant fields (NAME_2, ID_3, NAME_3, County, GDPPC):

```{r}
hunan_sf = hunan_sf %>% select(1:3, 7, 15)
```

Visualising the regional development indicator (GDP per Capita):

```{r}
equal = tm_shape(hunan_sf) +
  tm_fill("GDPPC", n = 5, style = "equal") +
  tm_borders(alpha = 1) +
  tm_layout(main.title = "Equal interval classification",
            main.title.size = 1)

quantile = tm_shape(hunan_sf) +
  tm_fill("GDPPC", n = 5, style = "quantile") +
  tm_borders(alpha = 1) +
  tm_layout(main.title = "Equal quantile classification",
            main.title.size = 1)

tmap_arrange(equal, quantile, ncol = 2)
```

## Global Measures of Spatial Autocorrelation

Compute Contiguity Spatial Weights using *poly2nb:*

```{r}
wm_q = poly2nb(hunan_sf, queen = TRUE)
summary(wm_q)
```

Compute the Row-standardised weights matrix:

```{r}
rswm_q = nb2listw(wm_q, 
                  style = "W", 
                  zero.policy = TRUE #Add weights of zero length for regions without neighbours
                  )
rswm_q
```

### Moran's I

Perform Moran's I statistics testing:

```{r}
moran.test(hunan_sf$GDPPC,
           listw = rswm_q,
           zero.policy = TRUE,
           na.action = na.omit)
#na.omit to remove all "NA"(s) in the data
```

A positive Moran I value is observed, suggesting potential clustering in GDPPC data. It could suggest that counties with similar GDPPC tend to be close to each other.

Perform Monte Carlo Moran's I:

```{r}
set.seed(1234)
bperm = moran.mc(hunan_sf$GDPPC,
                 listw = rswm_q,
                 nsim = 999,
                 zero.policy = TRUE,
                 na.action = na.omit)
bperm
```

As the p-value \< 0.05, we can reject the null hypothesis of complete spatial randomness (CSR).

Plotting and visualising the results:

```{r}
#| fig-height: 5
hist(bperm$res, freq = TRUE, breaks = 20, xlab = "Simulated Moran's I")
abline(v=0, col = "red")
```

### Geary's C

Perform Geary's C test:

```{r}
geary.test(hunan_sf$GDPPC, listw = rswm_q)
```

As the Geary C value is less than 1, we can draw the same conclusion as in the Moran's I test: There is signs of clustering between the counties' GDPPC values.

Perform Monte Carlo test for Geary's C:

```{r}
set.seed(1234)
bperm_geary = geary.mc(hunan_sf$GDPPC,
                       listw = rswm_q,
                       nsim = 999)
bperm_geary
```

Once again, with p-value \< 0.05, we can reject the null hypothesis of CSR.

Visualising the Monte Carlo results for Geary's C:

```{r}
#| fig-height: 5
hist(bperm_geary$res, freq = TRUE, breaks = 20, xlab = "Simulated Geary C")
abline(v=1, col = "red")
```

### Spatial Correlogram

Spatial correlograms show how correlated are pairs of spatial observations when you increase the distance(lag) between them.

Compute Moran's I correlogram:

```{r}
#| fig-height: 5
MI_corr = sp.correlogram(wm_q,
                         hunan_sf$GDPPC,
                         order = 6,
                         method = "I",
                         style = "W")
plot(MI_corr)
```

Plotting may not give a complete interpretation, as not all values are statistically significant. We must examine the full analysis report:

```{r}
print(MI_corr)
```

We can do the same for Geary's C:

```{r}
#| fig-width: 8
GC_corr = sp.correlogram(wm_q,
                         hunan_sf$GDPPC,
                         order = 6,
                         method = "C",
                         style = "W")
plot(GC_corr)
```

```{r}
print(GC_corr)
```

From both, we can see as we increase lag, we do see a shift of values towards the neutral (I = 0, C = 1), signifying the distribution of GDPPR to be closer to CSR with a large enough lag.

## Local Measures of Spatial Autocorrelation

### Local Indicators of Spatial Association (LISA)

Computing of Contiguity Spatial Weights and Row-Standardised Weight Matrix already done under Global Measures. We can use the same computation to compute the local Moran's I:

```{r}
fips = order(hunan_sf$County)
localMI = localmoran(hunan_sf$GDPPC, rswm_q)
head(localMI)
```

-   Ii: the local Moran’s I statistics

-   E.Ii: the expectation of local moran statistic under the randomisation hypothesis

-   Var.Ii: the variance of local moran statistic under the randomisation hypothesis

-   Z.Ii:the standard deviate of local moran statistic

-   Pr(): the p-value of local moran statistic

Use *printCoefmat()* to list the local Moran matrix:

```{r}
printCoefmat(data.frame(
  localMI[fips,],
  row.names = hunan_sf$County[fips]),
  check.names = FALSE)
```

Appending the local Moran's I onto the Hunan data frame:

```{r}
hunan.localMI = cbind(hunan_sf, localMI) %>%
  rename(Pr.Ii = Pr.z....E.Ii..)
```

Now we can map the local Moran's I values and the p-values using *tmap:*

```{r}
#| fig-width: 12

localMI.map = tm_shape(hunan.localMI) +
  tm_fill(col = "Ii",
          style = "pretty",
          palette = "GnBu",
          title = "Local Moran's I stats") +
  tm_borders(alpha = 1)

pvalue.map = tm_shape(hunan.localMI) +
  tm_fill(
    col = "Pr.Ii",
    breaks = c(-Inf, 0.001, 0.01, 0.05, 0.1, Inf),
    palette = "-Oranges",
    title = "Local Moran's I p-values"
  ) + tm_borders(alpha = 1)

tmap_arrange(localMI.map, pvalue.map, ncol = 2)
```

### Creating a LISA Cluster Map

Moran scatterplot shows the relationship between the values of each location and the average value at neighbouring locations:

```{r}
#| fig-width: 8
nci = moran.plot(hunan_sf$GDPPC, rswm_q,
                 labels = as.character(hunan_sf$County),
                 xlab = "GDPPC 2012",
                 ylab = "Spatially Lag GDPPC 2012")
```

The area is split into 4 quadrant: The top right is for areas that have high GDPPC and are surrounded by other areas that have average level of GDPPC.

Now to plot a Moran scatterplot with standardised variable, we first use *scale()* to center and scale the variable:

```{r}
#| fig-width: 8
hunan_sf$Z.GDPPC = scale(hunan_sf$GDPPC) %>% as.vector()
nci2 = moran.plot(hunan_sf$Z.GDPPC, rswm_q,
                  labels = as.character(hunan_sf$County),
                  xlab = "z-GDPPC 2012",
                  ylab = "Spatially Lag z-GDPPC 2012")
```

Now we will prepare the LISA map classes:

```{r}
quadrant = vector(mode = "numeric", length = nrow(localMI))

#Derive spatially lagged variable and center it around its mean
hunan_sf$lag_GDPPC = lag.listw(rswm_q, hunan_sf$GDPPC)
DV = hunan_sf$lag_GDPPC - mean(hunan_sf$lag_GDPPC)

#Center the local Moran's aroudn the mean
LM_I = localMI[,1] - mean(localMI[,1])

#Set statistical significance level
signif = 0.05

#define the low-low to high-high categories
quadrant[DV<0 & LM_I>0] = 1
quadrant[DV>0 & LM_I<0] = 2
quadrant[DV<0 & LM_I<0] = 3  
quadrant[DV>0 & LM_I>0] = 4

#place non-significant Moran in cat 0
quadrant[localMI[,5]>signif] = 0
```

Plot the LISA map:

```{r}
#| fig-width: 12
hunan.localMI$quadrant = quadrant
colours = c("#ffffff", "#2c7bb6", "#abd9e9", "#fdae61", "#d7191c")
clusters = c("insignificant", "low-low", "low-high", "high-low", "high-high")

LISAmap = tm_shape(hunan.localMI) +
  tm_fill(col = "quadrant",
          style = "cat",
          palette = colours[c(sort(unique(quadrant)))+1],
          labels = clusters[c(sort(unique(quadrant)))+1],
          popup.vars = c("")) +
  tm_borders(alpha = 1)

tmap_arrange(localMI.map, LISAmap, ncol = 2)
```

### Hot & Cold Spot Area Analysis

Derive the centroid coordinates using *st_centroid:*

```{r}
longitude = map_dbl(hunan_sf$geometry, ~st_centroid(.x)[[1]])
latitude = map_dbl(hunan_sf$geometry, ~st_centroid(.x)[[2]])
coords = cbind(longitude, latitude)
```

Determine the cut-off distance (upper limit):

```{r}
k1 = knn2nb(knearneigh(coords))
k1dist = unlist(nbdists(k1, coords, longlat = TRUE))
summary(k1dist)
```

The upper limit is 61.79km, so we can calculate the fixed distance as 62km:

```{r}
wm_d62 = dnearneigh(coords, 0, 62, longlat = TRUE)
wm_d62
```

```{r}
wm62_lw = nb2listw(wm_d62, style = 'B')
summary(wm62_lw)
```

Using adaptive distance to calculate another weight matrix:

```{r}
knn = knn2nb(knearneigh(coords, k = 8))
knn
```

The adaptive has been set to have the average number of k=8 neighbours. Covert this into spatial weights object:

```{r}
knn_lw = nb2listw(knn, style = 'B')
summary(knn_lw)
```

### Computing Gi Statistics

Using fixed distance:

```{r}
gi.fixed = localG(hunan_sf$GDPPC, wm62_lw)
gi.fixed
```

Gi statistics is represented as a Z-score, where greater values = greater intensity of clustering, and direction (positive/negative) indicates high/low clusters.

```{r}
hunan.gi = cbind(hunan_sf, as.matrix(gi.fixed)) %>%
  rename(gstat_fixed = as.matrix.gi.fixed.)
```

Now we can plot it out, next to the normal values:

```{r}
#| fig-width: 12
Gimap = tm_shape(hunan.gi) +
  tm_fill(col = "gstat_fixed",
          style = "pretty",
          palette = "-RdBu",
          title = "local Gi") +
  tm_borders(alpha = 1)

tmap_arrange(qtm(hunan_sf, "GDPPC"), Gimap, ncol = 2)
```

Using adaptive distance:

```{r}
#| fig-width: 12

gi.adaptive = localG(hunan_sf$GDPPC, knn_lw)
hunan.gi = cbind(hunan_sf, as.matrix(gi.adaptive)) %>%
  rename(gstat_adaptive = as.matrix.gi.adaptive.)

Gimap = tm_shape(hunan.gi) +
  tm_fill(col = "gstat_adaptive",
          style = "pretty",
          palette = "-RdBu",
          title = "local Gi") +
  tm_borders(alpha = 1)

tmap_arrange(qtm(hunan_sf, "GDPPC"), Gimap, ncol = 2)
```
