---
title: "In-Class Exercise 10"
author: "Dexter Wan"
date: "October 28, 2024"
date-modified: "last-modified"
execute: 
  eval: true
  echo: true
  freeze: true
---

## Description

Learning about building hedonic pricing models in R.

### Changelog

28 Oct 24: Completed In-Class Exercise 10.

## Importing Data and Packages

```{r}
pacman::p_load(olsrr, ggstatsplot, sf, tmap, tidyverse, performance, see ,sfdep)
```

```{r}
condo_resale = read.csv("data/aspatial/Condo_resale_2015.csv")
mpsz = read_rds("data/rds/mpsz.rds")
condo_resale_sf = read_rds("data/rds/condo_resale_sf.rds")
```

## Correlation Analysis - ggstatsplot methods

```{r}
#| fig-width: 12
#| fig-height: 12
ggcorrmat(condo_resale[, 5:23])
```

```{r}
condo_mlr = lm(formula = SELLING_PRICE ~ AREA_SQM + AGE + PROX_CBD 
               + PROX_CHILDCARE+ PROX_ELDERLYCARE + PROX_URA_GROWTH_AREA 
               + PROX_HAWKER_MARKET + PROX_KINDERGARTEN + PROX_MRT 
               + PROX_PARK + PROX_PRIMARY_SCH + PROX_TOP_PRIMARY_SCH 
               + PROX_SHOPPING_MALL + PROX_SUPERMARKET + PROX_BUS_STOP 
               + NO_Of_UNITS + FAMILY_FRIENDLY + FREEHOLD,
               data = condo_resale_sf)
summary(condo_mlr)
```

```{r}
ols_regress(condo_mlr)
```

```{r}
olsrr_condo = ols_regress(condo_mlr)
```

### Variable selection

```{r}
condo_fw_mlr = ols_step_forward_p(
  condo_mlr,
  p_val = 0.05,
  details = FALSE) #disable comprehensive report, just get basic report

condo_fw_mlr
```

```{r}
plot(condo_fw_mlr)
```

```{r}
condo_bw_mlr = ols_step_backward_p(
  condo_mlr,
  p_val = 0.05,
  details = FALSE) #disable comprehensive report, just get basic report

condo_bw_mlr

condo_both_mlr = ols_step_both_p(
  condo_mlr,
  p_val = 0.05,
  details = FALSE) #disable comprehensive report, just get basic report

condo_both_mlr
```

### Model selection

```{r}
metric = compare_performance(condo_mlr,
                             condo_fw_mlr$model,
                             condo_bw_mlr$model,
                             condo_both_mlr$model)
metric$Name = gsub(".*\\\\([a-zA-Z0-9_]+)\\\\, \\\\model\\\\.*", 
                   "\\1", metric$Name)
```

```{r}
plot(metric)
```

```{r}
#| fig-height: 12
ggcoefstats(condo_both_mlr$model, sort = "ascending")
```

## Checking for multicollinearity

```{r}
check_collinearity(condo_both_mlr$model)
```

```{r}
plot(check_collinearity(condo_both_mlr$model)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

### Assumption test

linearlity

```{r}
out = plot(check_model(condo_both_mlr$model, panel = FALSE))
out[[2]]
```

Normality

```{r}
plot(check_normality(condo_both_mlr$model))
```

Checking of outliers

```{r}
outliers = check_outliers(condo_both_mlr$model, method = "cook")
outliers
```

```{r}
plot(outliers)
```

## Visualising spatial non-stationary

```{r}
mlr_output = as.data.frame(condo_fw_mlr$model$residuals) %>%
  rename(`FW_MLR_RES`=`condo_fw_mlr$model$residuals`)
```

```{r}
condo_resale_sf = cbind(condo_resale_sf, mlr_output$FW_MLR_RES) %>%
  rename(`MLR_RES`=`mlr_output.FW_MLR_RES`)
```

```{r}
tmap_mode("view")
tm_shape(mpsz) + 
  tmap_options(check.and.fix = TRUE) +
  tm_polygons(alpha = 0.4) +
tm_shape(condo_resale_sf) +
  tm_dots(col = "MLR_RES",
          alpha = 0.6,
          style = "quantile")
tmap_mode("plot")
```

## Spatial stationary test

```{r}
condo_resale_sf = condo_resale_sf %>%
  mutate(nb = st_knn(geometry, k=6, longlat = FALSE),
         wt = st_weights(nb, style = "W"),
         .before = 1)
```

```{r}
global_moran_perm(condo_resale_sf$MLR_RES, 
                  condo_resale_sf$nb, 
                  condo_resale_sf$wt, 
                  alternative = "two.sided", 
                  nsim = 99)
```
